Spatiotemporal Multi-Resolution Approximation of the Amari Type Neural Field Model

### 2012-02-04 Multiresolution_framework/ltx/manuscript.tex
minor edit
minor edit to EM
minor edits

\section{IDE Neural Field Model}
\singlespacing
\begin{table*}[!t]
\begin{tabular}{|l|l|l|}
	\hline
	\textbf{Symbol} & \textbf{Quantity} & \textbf{Units}\\
	\hline
	\multicolumn{3}{|c|}{\emph{Domain and indices}}\\
	\hline
	$\Omega$ & spatial domain & n.a.\\
	$\mathbf{r}$ & spatial location & [mm, mm]\\
	$t$ & time & s\\
	\hline
	\multicolumn{3}{|c|}{\emph{Model}}\\
	\hline
    $y_t$ & observation & mV\\
    $v(\mathbf{r},t)$ & mean membrane potential & mV \\
	$f(v\left(\mathbf{r},t\right))$ & firing rate function & spikes.s$^{-1}$\\
	$e(\mathbf{r},t)$ & field disturbance, with covariance function $\eta(\mathbf r-\mathbf r')$ & mV\\
	$\boldsymbol\varepsilon_t$ & observation noise, with covariance matrix $\Sigma_\varepsilon$ & mV\\
	$m(\mathbf{r}_n)$ & observation kernel where $n$ is sensor index $n=1,..,n_y$ & n.a. \\
	$w(\mathbf{r})$ & spatial connectivity kernel & n.a.\\
	$h(t)$ & post-synaptic response kernel & mV\\
	$\tau$, $\xi$ & membrane time constant \& time constant parameter & s, n.a.\\
	$\delta(t)$ & Dirac-delta function & n.a.\\
	\hline
	\multicolumn{3}{|c|}{\emph{Reduced Model}} \\
	\hline
   	$\mathbf{x}_t$ & state vector at time $t$ & mV\\
   	$\mathbf{e}_t$ & state disturbance vector, with covariance $\Sigma_e$ & mV\\
   	$\mathbf{C}$ & observation matrix & n.a. \\
	\hline
	\multicolumn{3}{|c|}{\emph{Frequency Analysis}} \\
	\hline
	$\boldsymbol{\nu}$, $\boldsymbol{\nu}_c$ & spatial frequency and spatial cutoff frequency & cycles/mm \\
	$\sigma_{\nu}^2$ & variance of Fourier transformed Gaussian basis function & (cycles/mm)$^2$\\
	\hline
	\multicolumn{3}{|c|}{\emph{Estimation}} \\
	\hline
	$\hat{\mathbf{x}}_t^{f-}$, $\hat{\mathbf{x}}_t^f$ & forward prior and posterior state estimates & mV\\
	$\hat{\mathbf{x}}_t^{b}$ & backward posterior state estimates & mV\\
	$P^{f-}_t$, $P^f_t$  & forward prior and posterior covariance matrices & mV$^2$\\
	$P^b_t$ & backward posterior covariance matrices & mV$^2$\\
	$M_t$& cross-covariance matrix & mV$^2$\\
	$\mathcal K_{t} $ & filter gain & n.a.\\ 
	$\mathcal S_{t} $ & smoother gain & n.a.\\
	\hline
\end{tabular}
\caption{\textbf{Notation.} The first column of the table gives the symbols for all the notation used in the paper. The second column provides a brief description of the quantity, and the third column provides the SI units where relevant.}
\label{tab:Notation}
\end{table*}  
\doublespacing

@@

Substituting $\mathbf A( \boldsymbol\theta)$ from Eq.~\eqref{eq:A_theta} into Eq.~\eqref{eq:Qfunction} gives

@@

The three traces that form Eq.~\eqref{eq:MRA-QintermsofTraces} are dealt with in turn. To begin, the first trace is written in terms of element-wise summations

@@

Two different experiments were designed to test the framework. In the first one, 200 realizations of 1 second of data were used, where the field disturbance, $e_t$, and the observation noise, $\boldsymbol\varepsilon_t$, were regenerated each run. This way the net shape of the connectivity kernel could be formed by averaging over the parameters estimates from each realization. In the second experiment, a single data set was generated and the field estimations are compared for five state-space models, each accounting for a different level of approximation, i.e. $j=0$ to $j=4$.

For each of the experiments, the estimation was applied to the final 900~ms, allowing the forward model's dynamics to stabilize from the initial conditions. The initial state, parameters and the support of the connectivity kernel were unknown to the estimator. The observation noise was set to $\boldsymbol\Sigma_{\varepsilon}=0.1 \times \mathbf{I}_{n_y}$ and the disturbance covariance function was set to $\gamma(r-r') = \varphi_{3,-2}(r-r')$ (B-spline, $j=3$).  The firing rate slope, $\varsigma$ was set to $0.56~mV^{-1}$ \citep{Wendling2005}. The observation kernel was modeled using a B-spline function, with a width of 0.8~mm at half the maximum amplitude where the distance between adjacent sensors was $0.5$~mm, resulting into $n_y = 161$ observations. The choice of B-spline function to model the observation kernel is justified by the experiment performed in \citep{Freestone2011}. The spacing and bandwidth of the sensors allowed the full spatial bandwidth of the field to be observed.  With this number of observations, the estimation problem was well-posed, i.e. the number of states, $n_x < n_y$ at each level of the decomposition. This is because the scaling functions are orthogonal to the wavelets at the same and higher levels of decomposition, and  the wavelets are orthogonal to each other across different scales (see Eqs.~\eqref{eq:MRA-PsiPsiOrthogonality} and \eqref{eq:MRA-PhiPsiOrthogonality}). 

>>>>

\section{IDE Neural Field Model}
\singlespacing
\begin{table*}[!t]
\begin{tabular}{|l|l|l|}
	\hline
	\textbf{Symbol} & \textbf{Quantity} & \textbf{Units}\\
	\hline
	\multicolumn{3}{|c|}{\emph{Domain and indices}}\\
	\hline
	$\Omega$ & spatial domain & n.a.\\
	$\mathbf{r}$ & spatial location & [mm, mm]\\
	$t$ & time & s\\
	\hline
	\multicolumn{3}{|c|}{\emph{Model}}\\
	\hline
    $y_t$ & observation & mV\\
    $v(\mathbf{r},t)$ & mean membrane potential & mV \\
	$f(v\left(\mathbf{r},t\right))$ & firing rate function & spikes.s$^{-1}$\\
	$e(\mathbf{r},t)$ & field disturbance, with covariance function $\eta(\mathbf r-\mathbf r')$ & mV\\
	$\boldsymbol\varepsilon_t$ & observation noise, with covariance matrix $\Sigma_\varepsilon$ & mV\\
	$m(\mathbf{r}_n)$ & observation kernel where $n$ is sensor index $n=1,..,n_y$ & n.a. \\
	$w(\mathbf{r})$ & spatial connectivity kernel & n.a.\\
	$h(t)$ & post-synaptic response kernel & mV\\
	$\tau$, $\xi$ & membrane time constant \& time constant parameter & s, n.a.\\
	$\delta(t)$ & Dirac-delta function & n.a.\\
	\hline
	\multicolumn{3}{|c|}{\emph{Reduced Model}} \\
	\hline
   	$\mathbf{x}_t$ & state vector at time $t$ & mV\\
   	$\mathbf{e}_t$ & state disturbance vector, with covariance $\Sigma_e$ & mV\\ 
    $\mathbf{A}(\boldsymbol{\theta})$& transition matrix& n.a.\\
   	$\mathbf{C}$ & observation matrix & n.a. \\
	\hline
	\multicolumn{3}{|c|}{\emph{Frequency Analysis}} \\
	\hline
	$\boldsymbol{\nu}$, $\boldsymbol{\nu}_c$ & spatial frequency and spatial cutoff frequency & cycles/mm \\
	$\sigma_{\nu}^2$ & variance of Fourier transformed Gaussian basis function & (cycles/mm)$^2$\\
	\hline
	\multicolumn{3}{|c|}{\emph{Estimation}} \\
	\hline
	$\hat{\mathbf{x}}_t^{f-}$, $\hat{\mathbf{x}}_t^f$ & forward prior and posterior state estimates & mV\\
	$\hat{\mathbf{x}}_t^{b}$ & backward posterior state estimates & mV\\
	$P^{f-}_t$, $P^f_t$  & forward prior and posterior covariance matrices & mV$^2$\\
	$P^b_t$ & backward posterior covariance matrices & mV$^2$\\
	$M_t$& cross-covariance matrix & mV$^2$\\
	$\mathcal K_{t} $ & filter gain & n.a.\\ 
	$\mathcal S_{t} $ & smoother gain & n.a.\\
	\hline
\end{tabular}
\caption{\textbf{Notation.} The first column of the table gives the symbols for all the notation used in the paper. The second column provides a brief description of the quantity, and the third column provides the SI units where relevant.}
\label{tab:Notation}
\end{table*}  
\doublespacing

@@

The factor of 2 is included to simplify the derivation in the later steps. Substituting $\mathbf A( \boldsymbol\theta)$ from Eq.~\eqref{eq:A_theta} into Eq.~\eqref{eq:Qfunction} gives

@@

The rest of the derivation is to isolate $\boldsymbol\theta$ from the terms forming $\mathcal{Q}$. The three traces that form Eq.~\eqref{eq:MRA-QintermsofTraces} are dealt with in turn. To begin, the first trace is written in terms of element-wise summations

@@

Two different experiments were designed to test the framework. In the first one, 200 realizations of 1 second of data were used, where the field disturbance, $e_t$, and the observation noise, $\boldsymbol\varepsilon_t$, were regenerated each run. This way the net shape of the connectivity kernel could be formed by averaging over the parameters estimates from each realization, the frequency response of the observed field used to decide on the level... . In the second experiment, a single data set was generated and the field estimations are compared for five state-space models, each accounting upto a certain level of approximation, i.e. $j=0$ to $j=4$.

For each of the experiments, the estimation was applied to the final 900~ms, allowing the forward model's dynamics to stabilize from the initial conditions. The initial state, parameters and the support of the connectivity kernel were unknown to the estimator. The observation noise was set to $\boldsymbol\Sigma_{\varepsilon}=0.1 \times \mathbf{I}_{n_y}$ and the disturbance covariance function was set to $\eta(r-r') = \varphi_{3,-2}(r-r')$ (B-spline, $j=3$).  The firing rate slope, $\varsigma$ was set to $0.56~mV^{-1}$ \citep{Wendling2005}. The observation kernel was modeled using a B-spline function, with a width of 0.8~mm at half the maximum amplitude where the distance between adjacent sensors was $0.5$~mm, resulting into $n_y = 161$ observations. The choice of B-spline function to model the observation kernel is justified by the experiment performed in \citep{Freestone2011}. The spacing and bandwidth of the sensors allowed the full spatial bandwidth of the field to be observed.  With this number of observations, the estimation problem was well-posed, i.e. the number of states, $n_x < n_y$ at each level of the decomposition. This is because the scaling functions are orthogonal to the wavelets at the same and higher levels of decomposition, and  the wavelets are orthogonal to each other across different scales (see Eqs.~\eqref{eq:MRA-PsiPsiOrthogonality} and \eqref{eq:MRA-PhiPsiOrthogonality}). 
